Reads, QC and trimming

### 2024-04-04 episodes/quality-control.Rmd
Checked QC

```bash
cd /proj/g2020004/nobackup/3MK013/<username>
interactive -A uppmax2024-2-10 -M snowy -t 04:00:00
pwd
```

@@

``` bash
cd /proj/g2020004/nobackup/3MK013/<username>/molepi/data
```

@@

In total, it should take about 15-20 minutes for FastQC to run on all four of our FASTQ files. When the analysis completes, your prompt will return. So your screen will look something like this:

@@

```bash
$ mkdir ../results/fastqc_untrimmed_reads
$ mv *.zip ../results/fastqc_untrimmed_reads/
$ mv *.html ../results/fastqc_untrimmed_reads/
```

@@

```bash
$ cd ../results/fastqc_untrimmed_reads/
```

@@

Now we will run seqtk trimfq on our data. To begin, navigate to your `data` directory:

```bash
cd molepi/data/
```

@@

We've just successfully trimmed one of our FASTQ files! However, there is some bad news. `seqtk` can only operate on one sample at a time and we have more than one sample. The good news is that we can use a `for` loop to iterate through our sample files quickly!

@@

``` bash
ls
```

``` output
...
ERR029206_1.copy.fastq 
ERR029206_1.fastq.gz
ERR029206_1.fastq.gz_trim.fastq 
ERR029206_1_trim.fastq 
ERR029206_2.fastq.gz
ERR029206_2.fastq.gz_trim.fastq
GCF_000195955.2_ASM19595v2_genomic.fna
```

@@

``` bash
mkdir trimmed_fastq
mv *fastq.gz_trim* trimmed_fastq
cd trimmed_fastq
ls
```

>>>>

```bash
interactive -A uppmax2024-2-10 -M snowy -t 04:00:00
cd /proj/g2020004/nobackup/3MK013/<username>
pwd
```

@@

```bash
cd /proj/g2020004/nobackup/3MK013/<username>/molepi/data
```

@@

In total, it should take a couple minutes minutes for FastQC to run on our two FASTQ files. When the analysis completes, your prompt will return. So your screen will look something like this:

@@

```bash
mkdir ../results/fastqc_untrimmed_reads
mv *.zip ../results/fastqc_untrimmed_reads/
mv *.html ../results/fastqc_untrimmed_reads/
```

@@

```bash
cd ../results/fastqc_untrimmed_reads/
```

@@

Now we will run seqtk trimfq on our data. To begin, make sure you are still in your `data` directory:

```bash
pwd
```

@@

We've just successfully trimmed one of our FASTQ files! However, there is some bad news. `seqtk` can only operate on one sample at a time and we have more than one sample. The good news is that we can use a `for` loop to iterate through our sample files quickly! This will take a few minutes.

@@

``` bash
ls ERR029206*
```

``` output
...
ERR029206_1_fastqc.html     ERR029206_1.fastq.gz_trim.fastq  ERR029206_2_fastqc.zip
ERR029206_1_fastqc.zip	    ERR029206_1_trim.fastq		       ERR029206_2.fastq.gz
ERR029206_1.fastq.gz        ERR029206_2_fastqc.html	         ERR029206_2.fastq.gz_trim.fastq
```

@@

``` bash
mkdir trimmed_fastq
mv *fastq.gz_trim* trimmed_fastq
cd trimmed_fastq
ls ERR029206*
```
